#Analysis of the Impact of Severe Weather in the USA (1950-2011)
##Synopsis
Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

The events in the database start in the year 1950 and end in November 2011.

##Data Processing
###Read Data
```{r,cache=TRUE, message=FALSE}
#read data and select the target data to save memory
storm <- read.csv(bzfile("repdata-data-StormData.csv.bz2"), header = TRUE)
storm <- subset(storm, select = c(EVTYPE, FATALITIES : CROPDMGEXP))

#Change all letters into lower case
storm$EVTYPE <- tolower(storm$EVTYPE)
storm$CROPDMGEXP <- tolower(storm$CROPDMGEXP)
storm$PROPDMGEXP <- tolower(storm$PROPDMGEXP)

#unify the expression of each event type
storm$EVTYPE <- gsub("^flash flood.*", "flash flood", storm$EVTYPE)
storm$EVTYPE <- gsub("*.flash flood.*", "flash flood", storm$EVTYPE)
storm$EVTYPE <- gsub("^flood.*", "flood", storm$EVTYPE)
storm$EVTYPE <- gsub("^winter storm.*", "winter weather", storm$EVTYPE)
storm$EVTYPE <- gsub("winter mix", "winter weather", storm$EVTYPE)
storm$EVTYPE <- gsub("wintery mix", "winter weather", storm$EVTYPE)
storm$EVTYPE <- gsub("wintry mix", "winter weather", storm$EVTYPE)
storm$EVTYPE <- gsub("^wild fire.*", "wildfire", storm$EVTYPE)
storm$EVTYPE <- gsub("^wildfire.*", "wildfire", storm$EVTYPE)
storm$EVTYPE <- gsub("^wild/forest fire.*", "wildfire", storm$EVTYPE)
storm$EVTYPE <- gsub("^wind.*", "high wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^wnd.*", "high wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^tstm wind.*", "high wind", storm$EVTYPE)
storm$EVTYPE <- gsub("high winds.*", "high wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^waterspout.*", "waterspout", storm$EVTYPE)
storm$EVTYPE <- gsub("^water spout.*", "waterspout", storm$EVTYPE)
storm$EVTYPE <- gsub("^wayterspout.*", "waterspout", storm$EVTYPE)
storm$EVTYPE <- gsub("^wet.*", "heavy rain", storm$EVTYPE)
storm$EVTYPE <- gsub("^heavy rain.*", "heavy rain", storm$EVTYPE)
storm$EVTYPE <- gsub("^volcanic.*", "volcanic ash", storm$EVTYPE)
storm$EVTYPE <- gsub("^cloud.*", "funnel cloud", storm$EVTYPE)
storm$EVTYPE <- gsub("^urban.*", "flood", storm$EVTYPE)
storm$EVTYPE <- gsub("^thunderstorm.*", "thunderstorm wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^thundeerstorm.*", "thunderstorm wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^thunderstrom.*", "thunderstorm wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^thunerstorm.*", "thunderstorm wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^tornado.*", "hurricane", storm$EVTYPE)
storm$EVTYPE <- gsub("^hurricane.*", "hurricane", storm$EVTYPE)
storm$EVTYPE <- gsub("^tropical storm.*", "tropical storm", storm$EVTYPE)
storm$EVTYPE <- gsub("^marine.*.wind.*", "marine thunderstorm wind", storm$EVTYPE)
storm$EVTYPE <- gsub("^lightning.*", "lightning", storm$EVTYPE)
storm$EVTYPE <- gsub("^hail.*", "hail", storm$EVTYPE)
storm$EVTYPE <- gsub("^heat.*", "heat", storm$EVTYPE)
storm$EVTYPE <- gsub("*.heat.*", "heat", storm$EVTYPE)
storm$EVTYPE <- gsub("excessiveheat", "heat", storm$EVTYPE)
storm$EVTYPE <- gsub("^blizzard.*", "blizzard", storm$EVTYPE)
storm$EVTYPE <- gsub("^heavy snow.*", "heavy snow", storm$EVTYPE)
storm$EVTYPE <- gsub("*.surf.*", "high surf", storm$EVTYPE)
storm$EVTYPE <- gsub("^coastal.*", "coastal flood", storm$EVTYPE)
storm$EVTYPE <- gsub("^cold.*", "cold/wind chill", storm$EVTYPE)
storm$EVTYPE <- gsub("^dry.*", "drought", storm$EVTYPE)
```

###find out which events are the most harmful to population health
Fatalities
```{r, cache=TRUE, message=FALSE}
fatalities <- aggregate(FATALITIES ~ EVTYPE,storm, sum)
fatalities <- fatalities[order(fatalities$FATALITIES, decreasing = TRUE),]
```
Injuries
```{r, cache=TRUE, message=FALSE}
injuries <- aggregate(INJURIES ~ EVTYPE, storm, sum)
injuries <- injuries[order(injuries$INJURIES, decreasing = TRUE),]
```

###find out which events have the greatest economic consequences
Property
```{r, cache=TRUE, message=FALSE}
#modify the exponent
storm$PROPDMGEXP[storm$PROPDMGEXP == "b"] <- 10000000000
storm$PROPDMGEXP[storm$PROPDMGEXP == "m"] <- 1000000
storm$PROPDMGEXP[storm$PROPDMGEXP == "k"] <- 1000
storm$PROPDMGEXP[storm$PROPDMGEXP == "h"] <- 100
storm$PROPDMGEXP[storm$PROPDMGEXP %in% c("+","-")] <- 1
storm$PROPDMGEXP[storm$PROPDMGEXP == ""] <- 0
#calculate the total number of property damaged
storm$prop <- as.numeric(storm$PROPDMGEXP) * as.numeric(storm$PROPDMG)
prop <- aggregate(prop ~ EVTYPE, storm, sum)
prop <- prop[order(prop$prop, decreasing = TRUE),]
```

Crop
```{r, cache=TRUE, message=FALSE}
#modify the exponent
storm$CROPDMGEXP[storm$CROPDMGEXP == "b"] <- 10000000000
storm$CROPDMGEXP[storm$CROPDMGEXP == "m"] <- 1000000
storm$CROPDMGEXP[storm$CROPDMGEXP == "k"] <- 1000
storm$CROPDMGEXP[storm$CROPDMGEXP == "?"] <- 0
#calculate the total number of crop damaged
storm$crop <- as.numeric(storm$CROPDMG) * as.numeric(storm$CROPDMGEXP)
crop <- aggregate(crop ~ EVTYPE, storm, sum)
crop <- crop[order(crop$crop, decreasing = TRUE),]
```

##Results
###Harm to population health
```{r, cache=TRUE, message=FALSE}
fatalities <- head(fatalities)
injuries <- head(injuries)

library(ggplot2)
pfata <- ggplot(data = fatalities, 
                aes(reorder(EVTYPE, FATALITIES), y= FATALITIES, fill = FATALITIES)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  xlab("Event") + ylab("Total number of fatalities (1950-2011)") +
  theme(legend.position = "none")

pinju <- ggplot(data = injuries,
                aes(reorder(EVTYPE, INJURIES),
                    y=INJURIES, fill = INJURIES)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  xlab("Event") + ylab("Total number of injuries (1950-2011)") + 
  theme(legend.position = "none")

print(pfata)
print(pinju)
```

###Ecnomic conquences
```{r, message=FALSE}
crop <- head(crop,4)
prop <- head(prop,4)

par(mfrow = c(2,1))
barplot(crop$crop, col = "grey", 
        main = "Total number of crop damaged (1950-2011)",
        names.arg = crop$EVTYPE)
barplot(prop$prop, col = "grey",
        main = "Total number of property damaged (1950-2011)",
        names.arg = prop$EVTYPE)
```

##Conclusion
Hurricane led to most fatalities and injuries.

River led to most crop damaged.

Flood led to most property damaged.
