---
output: pdf_document
---
##Statistical Inferences: Project
#Exponential distribution and Central Limit Theorem
Author: Dongyang Li


##Overview
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

##Simulations
###Simulation of exponential distribution and show the histogram of sample means.
```{r, cache=TRUE}
set.seed(1000)
samplemean <- NULL
for (i in 1:1000){
  samplemean <- c(samplemean, mean(rexp(40, rate = 0.2)))
}
hist(samplemean, col = "grey", 
     main = "Histogram of sample means", 
     xlab = "sample means")
```

###Compare the sample mean and theoretical mean
Calculate the theoretical mean of the distribution
```{r, cache=TRUE}
1/0.2
```

Calculate the sample mean
```{r, cache=TRUE}
mean(samplemean)
```

We can see the sample mean is close to the theoretical mean.

###Compare the sample variance and the theoretical variance
Calculate the theoretical variance
```{r, cache=TRUE}
(1/0.2)^2/40
```

Calculate the sample variance
```{r, cache=TRUE}
var(samplemean)
```

We can see the sample variance is close the the theoretical mean.

###Compare the distribution of sample means and normal distribution
```{r, cache=TRUE, message=FALSE}
library(ggplot2)
p <- ggplot(data.frame(samplemean), aes(samplemean)) +
  geom_histogram(aes(y = ..density..),  
                 color = "grey", fill = "wheat") +
  labs(title = "Histogram of sample means", x = "Sample means") +
  stat_function(fun = dnorm, color = "red", size = 2,
                args = list(mean = 5, sd = sqrt((1/0.2)^2/40))) +
  stat_function(fun = dnorm, color = "green", size = 2,
                args = list(mean = mean(samplemean), sd = sqrt(var(samplemean))))
print(p)
```

We can see the histogram is close to normal distribution.

##Conclusion
When the sample size is large enough, Central Limit Theorem is satisified.