#Reproducible Research Project 1


##Loading the processing the data

###Load the necessary package
```{r, cache=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(lattice)
```

###Load the data and make it clean
```{r, cache=TRUE}
activity <- read.csv("repdata-data-activity/activity.csv", 
                     na.strings = "NA")
activity$date <- as.Date(activity$date)
head(activity)
```

##What is mean total number of steps taken per day?

###Calculate the total number of steps taken per day
```{r, cache=TRUE}
totalno <- aggregate(steps ~ date, data = activity, na.rm = TRUE, sum)
head(totalno)
```

###Make a histogram of the total number of steps taken each day
```{r, cache=TRUE}
hist(totalno$steps, col = "grey", breaks = 10, 
     xlab = "Total steps per day",
     main = "Histogram of total steps per day")
```

###Calculate the mean of the total number of steps taken per day
```{r, cache=TRUE}
mean(totalno$steps)
```

###Calculate the median of the total number of steps taken per day
```{r, cache=TRUE}
median(totalno$steps)
```

##What is the average daily activity pattern?

###Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
Calculate the average number of steps
```{r, cache=TRUE}
average <- activity%>%
  group_by(interval)%>%
  summarize(meansteps = mean(steps, na.rm = TRUE))
```

Make plot
```{r, cache=TRUE}
with(average,plot(average$interval, average$meansteps, 
                  type = "l", xlab = "5 min interval",
                  ylab = "Average steps",
                  main = "Daily activity pattern"))
```

###Find out which interval contains the maximum number of steps
```{r, cache=TRUE}
with(average, interval[which.max(meansteps)])
```

##Inputing missing values
###Calculate and report the total number of missing values in the dataset
Calculate and report the number of missing values in steps
```{r, cache=TRUE}
sum(is.na(activity$steps))
```

Calculate and report the number of missing values in date
```{r, cache=TRUE}
sum(is.na(activity$date))
```

Calculated and report the number of missing values in interval
```{r, cache=TRUE}
sum(is.na(activity$interval))
```

Calculate and report the total number of missing values
```{r, cache=TRUE}
sum(is.na(activity$steps)) + sum(is.na(activity$date)) + sum(is.na(activity$interval))
```

###Fill in all of the missing values in the dataset
Use the mean for that 5-minute interval
```{r, cache=TRUE}
newact <- inner_join(activity, average, by = "interval")
newact$steps <- ifelse(is.na(newact$steps), 
                       newact$meansteps, newact$steps)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r, cache=TRUE}
newact <- subset(newact, select = steps : interval)
head(newact)
```

###Make a histogram of the total number of steps taken each day
```{r, cache=TRUE}
totalno2 <- aggregate(steps ~ date, data = newact, sum)
hist(totalno2$steps, col = "grey", breaks = 10,
     xlab = "Total steps per day", 
     main = "Histogram of total steps per day")
```

###Calculate and report the mean and median total number of steps taken per day
Calculate the mean
```{r, cache=TRUE}
mean(totalno2$steps)
```

Calculate the median
```{r, cache=TRUE}
median(totalno2$steps)
```

Mean and median are similar to the estimates from the first part of the assignment.  
Imputing missing values dose not affect the estimates of the total daily number of steps a lot.

##Are there differences in activity patterns between weekdays and weekends?

###Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r, cache=TRUE}
newact$day <- weekdays(newact$date)
newact$day <- ifelse(newact$day %in% c("Saturday", "Sunday"),
                     "weekend", "weekday")
head(newact)
```

###Make a panel plot
```{r, cache=TRUE}
newact_mean <- newact%>%
  group_by(interval, day)%>%
  summarize(meansteps = mean(steps))
xyplot(meansteps ~ interval | day, data = newact_mean,
       type = "l", layout = c(1,2),
       xlab = "5 min interval",
       ylab = "Average stpes")
```

Activity patterns of weekdays and activity patterns of weekends are different to each other.
